---
title: "Credentials"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Options}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(councilR)
```

`{councilR}` relies on `{keyring}` to securely and consistently access user credentials across machines. 

`{keyring}` works by accessing your machine's credential storage (macOS Keychain or Windows Credential Store). 

```{r assign-keyring, eval=FALSE}
keyring::key_set_with_value(
  "councilR.uid",
  "mc\\myuuid"
)

keyring::key_set_with_value(
  "councilR.pwd",
  "password"
)
```

Its important to note that your Council password has to be reset every three months. Be sure to update your `{keyring}` entry regularly. 

FRED-Oracle has separate credentials. See `r ?fred_oracle_connection()` for more information. 

# Updating your machine from previous councilR versions

Prior to version 0.3.0, we used `getOption("councilR.uid")` and`getOption("councilR.pwd")`. You can assign the new keyring values directly with the code below.

```{r keyring-migration}
library(keyring)
# assign councilR.uid
keyring::key_set_with_value(
  service = "councilR.uid",
  password = getOption("councilR.uid")
)

# do keyring and getOption  match?
getOption("councilR.uid") == keyring::key_get("councilR.uid")

# assign councilR.pwd
keyring::key_set_with_value(
  service = "councilR.pwd",
  password = getOption("councilR.pwd")
)

# do keyring and getOption match?
getOption("councilR.pwd") == keyring::key_get("councilR.pwd")
```

