---
title: "Options"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Options}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(councilR)
```

`{councilR}` has a few `options()` that can help the package run more smoothly. 

```{r}
options(
  councilR.uid = "mc\\myuuid", # your user id
  councilR.pwd = keyring::key_get("MetC") # your password
)
```

If you'd like to have these options loaded when you start up RStudio, edit your `.Rprofile`. An easy way to accomplish this is with `usethis::edit_r_profile()`.

```{r}
usethis::edit_r_profile()
```

This will open your `.Rprofile`.

Then, add `councilR.uid` and `councilR.pwd` to the list of options. 
```{r, eval=FALSE}
options(
  ...,
  counciR.uid = "mc\\myuuid",
  councilR.pwd = keyring::key_get("MetC")
)
```

Its important to note that your Council password has to be reset every three months. Be sure to update your `.Rprofile` or `{keyring}` entry regularly. 
