---
title: "Spatial Helpers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Spatial_helpers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r chunk-opts, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(councilR)
library(arcgislayers)
```

`{councilR}` has a couple functions and snippets to help you access and manipulate spatial data. 

## Data access  

Some datasets are maintained on [Minnesota Geospatial Commons](https://gis.data.mn.gov/). These datasets generally mirror datasets available in the internal GISLibrary. 

You can access these datasets using the [`{arcgislayers}`](https://r.esri.com/arcgislayers/) package.

On Minnesota Geospatial Commons, navigate to "REST Services" and use the search function to find the dataset you want. Click the "Copy" icon for the given dataset to copy the Feature Service URL.

Some Feature Services may reference multiple layers. You can query them with `arc_open()`, then append the layer number to the URL.

```{r arcgislayers-list}
# returns list of layers in the Feature Service
arcgislayers::arc_open("https://arcgis.metc.state.mn.us/data1/rest/services/planning/Regional_Parks_Trails/FeatureServer/")

# Access a specific layer by appending the layer number to the URL. For example, to access layer 0, append "/0" to the URL.
arcgislayers::arc_open("https://arcgis.metc.state.mn.us/data1/rest/services/planning/Regional_Parks_Trails/FeatureServer/0")
```

To pull a given layer, pass the URL to `arc_open()`, then use `arc_select()`. You can filter the data using the `where` argument. The `where` argument takes a SQL-like query string.

```{r regional-parks-commons}
regional_parks_commons <- arcgislayers::arc_open("https://arcgis.metc.state.mn.us/data1/rest/services/planning/Regional_Parks_Trails/FeatureServer/0") %>%
  arcgislayers::arc_select(
    where = "OMFundingAgency = 'Anoka County'"
  )

head(regional_parks_commons)
```

If you can access internal databases, use your Council login user ID and password to access data from the GIS Library. For seamless access, see `vignette("Credentials")` to learn how to set up your credentials.

```{r gis-query, eval = FALSE}
regional_parks_db <- import_from_gis(query = "GISCD.dbo.RPTAdministrativeBoundary")

head(regional_parks_db)
```

## Other helpers
A commonly fetched spatial dataset is the 7-county geography. `fetch_county_geo()` does just that.  
You can also make a map quickly by using `map_council_continuous()`. This also adds a scale bar and cardinal indicator. 

```{r plot-county, eval = TRUE}
fetch_county_geo(progress_bar = FALSE) %>%
  map_council_continuous(.fill = ALAND) +
  ggplot2::labs(title = "Counties")
```


You can also fetch CTUs. 

```{r plot-city, eval = TRUE}
fetch_ctu_geo(progress_bar = FALSE) %>%
  map_council_continuous(.fill = ALAND) +
  ggplot2::labs(title = "CTUs")
```


## Snippets

Snippets are text macros that are used for quickly inserting snippets of code in RStudio.

Once you have installed snippets using `councilR::snippets_install()`, you will now have important values quickly accessible. To access them, start typing the snippet name into the console. The snippet will autocomplete. 

  - `metro_crs`, the numeric preferred coordinate reference system (CRS) for our region. Most commonly used when transforming a spatial object's projection with `sf::st_transform(df, 26915)`
  - `tile_crs`, numeric coordinate reference system for spatial data using a tile-based system, like `{leaflet}`. Most commonly used with `sf::st_transform(df, 4326)`
  - `metro_centroid`, vector with latitude and longitude coordinates of the geographic center of the 7-county metro. Useful for setting a starting center point and zoom in `{leaflet}` maps.
  - `counties`, vector of county  names for the seven core counties serviced by the Metropolitan Council 
  - `counties_mpo`, county names for all counties included in the Metropolitan Planning Organization (MPO) area. 

```{r snippets, eval=FALSE}
metro_crs
#> 26915

metro_centroid
#> c(-93.30375,
#>   44.91831)
```
